# Loading in the data
```{r setup, include=FALSE}
library(readr)
EPI_data <- read_csv("epi2024results06022024.csv")
attach(EPI_data)
```

# Removing NA Values
```{r}
tf <- is.na(EPI.new) # records True values if the value is NA
EPI.new.no_NA <- EPI.new[!tf] # filters out NA values, new array 
```

# Exploring the distribution
```{r}
summary(EPI.new) # stats
fivenum(EPI.new,na.rm=TRUE)
stem(EPI.new) # stem and leaf plot
hist(EPI.new)
hist(EPI.new, seq(20., 80., 1.0), prob=TRUE)
lines(density(EPI.new,na.rm=TRUE,bw=1.)) # or try bw=“SJ”
rug(EPI.new)
```

# Creating a Box Plot of the Data
```{r}
boxplot(EPI.new, APO.new) 
```

# More Detailed Histogram
```{r}
hist(EPI.new, seq(20., 80., 1.0), prob=TRUE)
lines(density(EPI.new,na.rm=TRUE,bw=1.))
rug(EPI.new)
```


# More Detailed Histogram
```{r}
hist(EPI.new, seq(20., 80., 1.0), prob=TRUE)
lines(density(EPI.new,na.rm=TRUE,bw="SJ"))
rug(EPI.new)
```

# More Detailed Histogram
```{r}
hist(EPI.new, seq(20., 80., 1.0), prob=TRUE)
x<-seq(20,80,1)
q<-dnorm(x,mean=42, sd=5,log=FALSE)
lines(x,.4*q)
q<-dnorm(x,mean=65, sd=5,log=FALSE)
lines(x,.12*q) 
```

## Excercise 2: Fitting a Distribution Beyond Histograms

# Cumulative Density Function
```{r}
plot(ecdf(EPI.new), do.points=FALSE, verticals=TRUE)
```

# Quantile-Quantile
```{r}
qqnorm(EPI.new); qqline(EPI.new)
```

# Quantile-Quantile Plot Against Normal and T Distributions
```{r}
qqplot(rnorm(250), EPI.new, xlab = "Q-Q plot for norm dsn")
qqline(EPI.new)

qqplot(rt(250, df = 5), EPI.new, xlab = "Q-Q plot for t dsn")
qqline(EPI.new)
```

# Excercise 2a: Fitting a Distribution for variable APO
```{r}
# cumulative density function
plot(ecdf(APO.new), do.points=FALSE, verticals=TRUE)

# Quantile-Quantile plot
qqnorm(APO.new); qqline(APO.new)

# Quantile plot against normal distribution
qqplot(rnorm(250), APO.new, xlab = "Q-Q plot for norm dsn")
qqline(APO.new)

# Quantile plot against T distribution
qqplot(rt(250, df = 5), APO.new, xlab = "Q-Q plot for t dsn")
qqline(APO.new)
```

# Excercise 2a: Fitting a Distribution for variable APOe
```{r}
# cumulative density function
plot(ecdf(MHP.new), do.points=FALSE, verticals=TRUE)

# Quantile-Quantile plot
qqnorm(MHP.new); qqline(MHP.new)

# Quantile plot against normal distribution
qqplot(rnorm(250), MHP.new, xlab = "Q-Q plot for norm dsn")
qqline(MHP.new)

# Quantile plot against T distribution
qqplot(rt(250, df = 5), MHP.new, xlab = "Q-Q plot for t dsn")
qqline(MHP.new)
```

# Quantile-Quantile
```{r}
qqnorm(EPI.new); qqline(EPI.new)
```

# Quantile-Quantile
```{r}
qqnorm(EPI.new); qqline(EPI.new)
```

# Quantile-Quantile
```{r}
qqnorm(EPI.new); qqline(EPI.new)
```
